---
title: "3D Earth-tryout1"
author: "Ruxue Peng, rp2815"
date: "February 14, 2017"
output: html_document
---  
#Step 0: Install Required packages  
```{r,echo = F}
#set your working directory into the project folder and inside the lib library
setwd("F:/second_term/ADS/proj2/Spr2017-proj2-grp11-master/data")
packages.used <- 
  c("geosphere", # For spatial methods  
    "threejs",   # threejs is used for 3-D interactive Earth Visualization
    "rworldmap", # For creating earth map
    "leaflet",   # Leaflet for R provides functions to control and integrate Leaflet, a JavaScript library for interactive maps, within R.
    "rgeos",      # Provides functions for handling operations on topologies.
    "raster",     # For raster image
    "DT",         # For creating interactive tables
    "ggplot2",
    "sp"   ,       # For Spatial processing of data
    "ggmap",       # To reverse geocode Long/Lat
    "knitr",        # TO enable 3-D visualization embedding in the HTML page
    "rglwidget",
    "rgl"
    )
#
# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

#load the packages
library("geosphere")
library("threejs")
library("rworldmap")
library("leaflet")
library("rgeos")
library("raster")
library("DT")
library("ggplot2")
library("sp")
library("ggmap")
library("knitr")
library("rglwidget")
library("rgl")
source("../lib/Match_flora.R")
#
# Set this property to enable 3-D visualization
#
knit_hooks$set(webgl = hook_webgl)
```  

#Step 1: (First-time-only)Data Cleaning  
*skip if you have import_long_lat.csv in your data folder.  

```{r Data Cleaning}
# Read the file
if(FALSE){ 
setwd("F:/second_term/ADS/proj2/Spr2017-proj2-grp11-master/data")
input_data<-read.csv("data(cleaned).csv", header = TRUE, stringsAsFactors = FALSE)
str(input_data)
#$ Year                       : int  1996 1996 1996 1996 1996 1996 1996 1996 1996 1996 ...
# $ Country                    : chr  "Canada" "Mexico" "Guatemala" "Belize" ...
# $ ExportsFASValueBasisYtdDec : int  532850043 56077722 1656512 350102 1114079 1226421 642139 2113706 4827307 1283338 ...
# $ GenImportsCIFValBasisYtdDec: num  3.59e+08 6.32e+08 2.69e+08 1.33e+05 5.54e+07 ...
}
```  

#Step 2: (First-time-only)geocode to generate longitude and latitude data  
*skip if you have import_long_lat.csv in your data folder.  

```{r Geocoding}
#Make sure you have internet since geocode function uses google 
#get the longitue and latitude for the 46 countries in the dataset
if(FALSE){
long_lat = matrix(NA,nrow = length(unique(input_data$Country)),ncol = 2)
long_lat <- geocode(unique(input_data$Country))
long_lat <- cbind(unique(input_data$Country),long_lat)
names(long_lat) <- c("Country","Longitude","Latitude")
#match the 227 countries with the 3616 import records 
long.lat = apply(input_data,1,Match.flora)
long.lat.matrix = matrix(long.lat,ncol = 2,byrow = T)
input_data = cbind(input_data,long.lat.matrix)
setwd("F:/second_term/ADS/proj2/Spr2017-proj2-grp11-master/data")
write.csv(input_data,file = "import_long_lat.csv",row.names = F)
}

```  

#Step 2: loading geocoded csv from data folder  

```{r}
setwd("F:/second_term/ADS/proj2/Spr2017-proj2-grp11-master/data")
input_data =  read.csv("import_long_lat.csv",header = T,as.is = T)
names(input_data) <- c("Year","Country","export value",
                       "import value","longitude","latitude")
input_data=input_data[,1:6]
datatable(input_data,rownames = F)
#Can search for different countries' data
```  

#Step 3: subset the data we want- for shiny's input

```{r preparation for shiny input}
year = 1996
type = "import value"
input_data = input_data[input_data$Year == year,]
```


#Step 4: Converting data to Spatial Format
```{r Spatial Format}
# Converting Data to spatial objects
source_df<-data.frame(SourceLong=input_data$longitude,SourceLat=input_data$latitude)
# Create object of class SpatialPoints using SpatialPoints function from sp package
source_sp<-SpatialPoints(source_df, proj4string=CRS("+proj=longlat"))
str(source_sp)
head(source_sp)

# Convert to Spatial Dataframe
source_spdf <- SpatialPointsDataFrame(source_sp, data = source_df)
str(source_spdf)
head(source_spdf)
```  



```{r US long and lat}
# Similarly for the Destination Locations-US
C <- unlist(c(geocode("United States")[1],geocode("United States")[2]))
US <- data.frame(US_longitude= rep(C[1],length(input_data$`Imported From`)),
                US_latitude = rep(C[2]),length(input_data$`Imported From`))[,1:2]
# Converting Data to spatial objects

dest_df<-data.frame(DestLong=US$US_longitude,DestLat=US$US_latitude)
#
# Create object of class SpatialPoints using SpatialPoints function from sp package
#
dest_sp<-SpatialPoints(dest_df, proj4string=CRS("+proj=longlat"))
#
# Convert to Spatial Dataframe
#
dest_spdf <- SpatialPointsDataFrame(dest_sp, data = dest_df)
head(dest_spdf)
```  

#Step 5: Calculate the distance and bearing between import country and US  
*Bearing: A bearing is the angular difference away from a north or south baseline, ranging from 0° to 90°. Determining your bearing begins with a north or south baseline, whichever is closer to your direction. The angle is then measured east or west from the baseline.
```{r}
#using the  distHaversine and bearing functions from geosphere package
#Create a Combined data frame for display purpose only
comb_df<-cbind(input_data,To = rep("US",length(input_data$`Imported From`)),US)

# Calculate distance between Source and Destination
comb_df$distance<-distHaversine(source_sp,dest_sp)
comb_df$bearing<-bearing(dest_sp,source_sp)

# Display the combined dataframe using interactive table 
datatable(comb_df, rownames = FALSE)
```  

#Step 6: 2D without the map
```{r}
#
# Plot source points
#
plot(comb_df[5:6],col= "red", pch = 3,xlab="",ylab="")
#
# Add Destination points
#
points(comb_df[8:9],col="blue")

```  

#Step 7: with the map

```{r}
(worldMap <- getMap())
world.points <- fortify(worldMap)  # Convert data into dataframe using fortify from ggplot2
head(world.points)
#
world.points$region <- world.points$id
#
world.df <- world.points[,c("long","lat","group", "region")]
#
worldmap <- ggplot() + 
  geom_polygon(data = world.df, aes(x = long, y = lat, group = group)) +
  geom_point(aes(x=comb_df[,5], y=comb_df[,6]),color="yellow", size=1) +  # Plot Source Locations
  geom_point(aes(x=comb_df[,8], y=comb_df[,9]),color="cornflowerblue", size=10) +  # Plot Dest Location
  scale_y_continuous(breaks = (-2:2) * 30) +
  scale_x_continuous(breaks = (-4:4) * 45) +
  theme_bw() +
  theme(axis.line = element_line(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())
#
worldmap
```  

#Step 8: 3D map

```{r}
worldmap +coord_map("ortho", 
   orientation=c(40, -40,-10))+
  theme(panel.background = element_rect(fill = 'gray', colour = 'white'), panel.grid.major = element_line(color = "white"),
    panel.grid.minor = element_line(color = "white"))
```

#Step 9: add inactivity  

```{r}
comb_df$`import value` = comb_df$`import value`+1
comb_df$log = log(comb_df$`import value`) * 0.1
```

```{r}
(earth <- system.file("images/world.jpg",  package="threejs")) # Get the image of globe
#
# 3-D visual Dataframe
globejs(img=earth, lat=comb_df[,6], long=comb_df[,5],value = 1,color = "yellow", arcs=comb_df[,c(6,5,9,8)],
        arcsHeight=0.3, arcsLwd=comb_df$log, arcsColor="cornflowerblue", arcsOpacity=comb_df$log/max(comb_df$log)+0.5,
        atmosphere=TRUE, height=800, width = 800)
```

