---
title: "R Notebook"
output: html_notebook
---
#Step 0: Install Required packages  
```{r,echo = F,message=F}
#before everything, first set your working directory to the data folder
packages.used <- 
  c("geosphere", # For spatial methods  
    "threejs",   # threejs is used for 3-D interactive Earth Visualization
    "rworldmap", # For creating earth map
    "leaflet",   # Leaflet for R provides functions to control and integrate Leaflet, a JavaScript library for interactive maps, within R.
    "rgeos",      # Provides functions for handling operations on topologies.
    "raster",     # For raster image
    "DT",         # For creating interactive tables
    "ggplot2",
    "sp"   ,       # For Spatial processing of data
    "ggmap",       # To reverse geocode Long/Lat
    "knitr",        # TO enable 3-D visualization embedding in the HTML page
    "rglwidget",
    "rgl",
    "plyr",
    "reshape2"
    )
#
# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

#load the packages
library("plyr")
library("reshape2")
library("geosphere")
library("threejs")
library("rworldmap")
library("leaflet")
library("rgeos")
library("raster")
library("DT")
library("ggplot2")
library("sp")
library("ggmap")
library("knitr")
library("rglwidget")
library("rgl")
source("../lib/Match_flora.R")
source("../lib/plot_3D.R")

#
# Set this property to enable 3-D visualization
#
knit_hooks$set(webgl = hook_webgl)
```  

#Step 1: loading geocoded csv from data folder  

```{r}
input_data =  read.csv("import_long_lat.csv",header = T,as.is = T)
names(input_data) <- c("Year","Country","export value",
                       "import value","longitude","latitude")
input_data=input_data[,1:6]
#datatable(input_data,rownames = F)
#Can search for different countries' data
```  

#Step 2: subset the data we want- for shiny's input

```{r preparation for shiny input}
input_data = melt(input_data,id.vars = c("Year","Country","longitude","latitude"),
                  variable.name = "type")
year = 1996
type = "import value"
input_data = input_data[input_data$Year == year & input_data$type == type,]
num_of_arc = 100
input_data = input_data[order(input_data$value,decreasing = T),]
input_data = input_data[1:num_of_arc,]
```


#Step 3: Converting data to Spatial Format
```{r Spatial Format}
# Converting Data to spatial objects
source_df<-data.frame(SourceLong=input_data$longitude,SourceLat=input_data$latitude)
# Create object of class SpatialPoints using SpatialPoints function from sp package
source_sp<-SpatialPoints(source_df, proj4string=CRS("+proj=longlat"))
str(source_sp)
head(source_sp)

# Convert to Spatial Dataframe
source_spdf <- SpatialPointsDataFrame(source_sp, data = source_df)
str(source_spdf)
head(source_spdf)

# Similarly for the Destination Locations-US
C <- unlist(c(geocode("United States")[1],geocode("United States")[2]))
US <- data.frame(US_longitude= rep(C[1],length(input_data$`Country`)),
                US_latitude = rep(C[2]),length(input_data$`Country`))[,1:2]
# Converting Data to spatial objects

dest_df<-data.frame(DestLong=US$US_longitude,DestLat=US$US_latitude)
#
# Create object of class SpatialPoints using SpatialPoints function from sp package
#
dest_sp<-SpatialPoints(dest_df, proj4string=CRS("+proj=longlat"))
#
# Convert to Spatial Dataframe
#
dest_spdf <- SpatialPointsDataFrame(dest_sp, data = dest_df)
```  


#Step 4: Calculate the distance and bearing between import country and US  
*Bearing: A bearing is the angular difference away from a north or south baseline, ranging from 0° to 90°. Determining your bearing begins with a north or south baseline, whichever is closer to your direction. The angle is then measured east or west from the baseline.
```{r}
#using the  distHaversine and bearing functions from geosphere package
#Create a Combined data frame for display purpose only
comb_df<-cbind(input_data,To = rep("US",length(input_data$`Country`)),US)

# Calculate distance between Source and Destination
comb_df$distance<-distHaversine(source_sp,dest_sp)
comb_df$bearing<-bearing(dest_sp,source_sp)


```

```{r}
comb_df$`value` = comb_df$`value`+1
comb_df$log = log(comb_df$`value`) * 0.1
```  

#Step 7: (plan B)

```{r}
library("maptools")
data(wrld_simpl)                             # Basic country shapes
load(url("http://illposed.net/bono.rdata"))  # Awareness index data

bgcolor <- "#000025"
earth <- tempfile(fileext=".jpg")
jpeg(earth, width=2048, height=1024, quality=100, bg=bgcolor, antialias="default")
par(mar = c(0,0,0,0),    pin = c(4,2),    pty = "m",    xaxs = "i",
    xaxt = "n",          xpd = FALSE,    yaxs = "i",    yaxt = "n")

map_palette <- apply(col2rgb(heat.colors(5)[5:1])/768,2,function(x)rgb(x[1],x[2],x[3]))
# Restrict bono data to countries from the maptools package
bono <- bono[rownames(bono) %in% wrld_simpl$NAME, ,drop=FALSE]
# Set a default color for each country and the colors from the bono data
clrs <- rep(map_palette[1], length(wrld_simpl$NAME))
names(clrs) <- wrld_simpl$NAME
clrs[rownames(bono)] <- map_palette[bono$index]

plot(wrld_simpl,  col=clrs,   bg=bgcolor,        border="cyan",  ann=FALSE,
     axes=FALSE,  xpd=FALSE,  xlim=c(-180,180), ylim=c(-90,90),  setParUsrBB=TRUE)

graphics.off()

legendcol=heat.colors(5)[5:1]
globejs(img=earth, bg="white",
        arcs=comb_df[,c(4,3,9,8)],
        arcsHeight=0.4, 
        arcsLwd=comb_df$log,
        arcsColor="cornflowerblue", 
        arcsOpacity=0.8,
        atmosphere=TRUE, height=600, width = 600
        )

```

